Manuscript draft

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig. 4), with no major trends in C allocatio (Fig. 2). This finding contradicts some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming. Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$). The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**). The latter understanding is driven by observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; @anderson-teixeira_carbon_2016] (**REFS(KAT)**), but our analysis suggests that these low values may more appropriately be attributed to the fact that these forests are old than to their climate. Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**). Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.  [There is some indication that $GPP$ may have a somewhat steeper and less variable relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig. 2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**). (**Helene, maybe comment on Larjavaara & Muller-Landau (2011)?**). ] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients. 
[There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006]. That we do not see consistent patterns of variation in allocation across global climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales and/or in response to factors other than mean climate, including inter- and intra-annual climiatic fluctuations. ]

One interesting observation was that climate explains more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig. 2; **some SI table?**). There are two non-exclusive potential explanations for this. First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$**), and sampling depth is variable and often insufficient to capture the full soil profile. $ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**). However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes. Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes. This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes. [**work on this second part, add refs**]

>>>>

[**THIS PARAGRAPH NEEDS ATTENTION!**]. We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig. 4), with no major trends in C allocation among the variable pairs tested (Fig. 2; **Some SI table**). *This finding does not support some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming.* Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$). The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**). The latter understanding is driven by observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; @anderson-teixeira_carbon_2016] (**REFS(KAT)**), but our analysis suggests that these low values may more appropriately be attributed to the fact that these forests are old than to their climate. Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**). Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.  [There is some indication that $GPP$ may have a somewhat steeper and less variable relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig. 2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**). (**Helene, maybe comment on Larjavaara & Muller-Landau (2011)?**). ] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients. 
[There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006]. That we do not see consistent patterns of variation in allocation across global climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales and/or in response to factors other than mean climate, including inter- and intra-annual climiatic fluctuations. ]

One interesting observation was that climate explains more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig. 2; **some SI table?**). There are two non-exclusive potential explanations for this. First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$**), and sampling depth is variable and often insufficient to capture the full soil profile. $ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**?- Becky, did we end up lumping? REF**). However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes. Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes. *This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.* [**work on this second part, add refs**]
